---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# refsampr

<!-- badges: start -->
[![R-CMD-check](https://github.com/AleSR13/refsampr/workflows/R-CMD-check/badge.svg)](https://github.com/AleSR13/refsampr/actions)
<!-- badges: end -->

The goal of refsamp is to make plots for tracking the quality of reference samples in a sequencing facility (WGS). It is tailor-made for the [RIVM](https://www.rivm.nl/), especially to be used directly on the output from the [Juno pipeline](https://github.com/AleSR13/Juno_pipeline) (provided that the date in the format 'yymmdd' is included in the outpur directory from Juno). The package can be expanded for external use.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("AleSR13/refsampr")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(refsamp)

make_refsamp_report("210120_Juno_pipeline_results",
                    output_report = "out/report.html",
                    history_data = "history_data.csv" )

```
